partial: An R Package for Creating Partial Dependence Plots

### 2016-08-09 manuscript/greenwell.tex
minor edits

The \code{partial} function relies on the \CRANpkg{plyr} package \citep{plyr-pkg}, rather than for loops. This makes it possible to request progress bars (e.g., \code{.progress = "text"}) or run \code{partial} in parallel. In fact, \code{partial} can use any of the parallel backends supported by the \pkg{foreach} package. To use this functionality, we must load and register a supported parallel backend (e.g., \CRANpkg{doMC} \citep{doMC-pkg} or \CRANpkg{doParallel} \citep{doParallel-pkg}). The following snippet of code obtains the partial dependence of median home value on \code{rm}, \code{lstat}, and \code{ptratio} in parallel. The result is displayed in Figure~\ref{fig:partial_par}.

Setting up a parallel backend is rather straightforward. To illustrate, the following snippet of sets up the \code{partial} function to run in parallel on Unix-like systems\footnote{This example will not run on Windows.} using the \pkg{doParallel}.
\begin{example}
library(doParallel)  # load parallel backend
registerDoParallel(cores = 4)  # use 4 cores
\end{example}
Now, to run \code{partial} in parallel, all we have to do is invoke the \code{.parallel = TRUE} option and the rest is taken care of by the internal call to \pkg{plyr} the parallel backend we loaded:
\begin{example}
pd <- partial(boston.rf, pred.var = c("lstat", "rm", "ptratio"),
              grid.resolution = 20, .parallel = TRUE)
plotPartial(pd, number = 4, overlap = 0.1)
\end{example}

It is important to note that when using more than two variables, \code{plotPartial} produces a trellis display. The first two variables given to \code{pred.var} are used for the horizontal and vertical axes; additional variables define the panels. If the panel variables are continuous, then shingles\footnote{A shingle is a special Trellis data structure that consists of a numeric vector along with intervals that define the "levels" of the shingle. The intervals may be allowed to overlap.} are produced first using the equal count algorithm. Hence, it is probably better to use categorical variables to define the panels in higher dimensional displays when possible.

>>>>

The \code{partial} function relies on the \CRANpkg{plyr} package \citep{plyr-pkg}, rather than for loops. This makes it possible to request progress bars (e.g., \code{.progress = "text"}) or run \code{partial} in parallel. In fact, \code{partial} can use any of the parallel backends supported by the \pkg{foreach} package. To use this functionality, we must load and register a supported parallel backend [e.g., \CRANpkg{doMC} \citep{doMC-pkg} or \CRANpkg{doParallel} \citep{doParallel-pkg}].

Setting up a parallel backend is rather straightforward. To illustrate, the following snippet of sets up the \code{partial} function to run in parallel on Unix-like systems\footnote{This example will not run on Windows.} using the \pkg{doParallel}.
\begin{example}
library(doParallel)  # load parallel backend
registerDoParallel(cores = 4)  # use 4 cores
\end{example}
Now, to run \code{partial} in parallel, all we have to do is invoke the \code{.parallel = TRUE} option and the rest is taken care of by the internal call to \pkg{plyr} the parallel backend we loaded. This is illustrated in following snippet of code which obtains the partial dependence of median home value on $\boldsymbol{z}_s = \left\{\code{rm}, \code{ptratio}, \code{chas}\right\} in parallel. The result is displayed in Figure~\ref{fig:partial_par}.
\begin{example}
pd <- partial(boston.rf, pred.var = c("rm", "ptratio", "chas"),
              grid.resolution = 20, .parallel = TRUE)
plotPartial(pd, number = 4, overlap = 0.1)
\end{example}

%Now, to run \code{partial} in parallel, all we have to do is invoke the \code{.parallel = TRUE} option and the rest is taken care of by the internal call to \pkg{plyr} the parallel backend we loaded. This is illustrated in following snippet of code which obtains the partial dependence of median home value on $\boldsymbol{z}_s = \left\{\code{rm}, \code{lstat}, \code{ptratio}\right\} in parallel. The result is displayed in Figure~\ref{fig:partial_par}.
%\begin{example}
%pd <- partial(boston.rf, pred.var = c("lstat", "rm", "ptratio"),
%              grid.resolution = 20, .parallel = TRUE)
%plotPartial(pd, number = 4, overlap = 0.1)
%\end{example}

It is important to note that when using more than two variables, \code{plotPartial} produces a trellis display. The first two variables given to \code{pred.var} are used for the horizontal and vertical axes; additional variables define the panels. If the panel variables are continuous, then shingles\footnote{A shingle is a special Trellis data structure that consists of a numeric vector along with intervals that define the "levels" of the shingle. The intervals may be allowed to overlap.} are produced first using the equal count algorithm. Hence, it is probably better to use categorical variables to define the panels in higher dimensional displays when possible, as was done in the previous example.
